Evidence accumulation modelling offers new insights into the cognitive mechanisms that underlie linguistic and action-based training

### 2025-04-01 manuscript/manuscript.Rmd
Minor updates

---
title             : "Evidence accumulation modelling offers new insights into the cognitive mechanisms that underlie linguistic and action-based training"
shorttitle        : "Decision models and experience-dependent plasticity"
author: 
  - name          : "Samantha Parker"
    affiliation   : "1"
    # corresponding : yes    # Define only one corresponding author
    # address       : "School of Psychological Sciences, 16 University Avenue, Macquarie University, NSW 2109"
    # email         : "samantha.parker@students.mq.edu.au"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Data curation"
      - "Formal analysis"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Emily S. Cross"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Resources"
      - "Writing - Review & Editing"
  - name          : "Richard Ramsey"
    affiliation   : "2"
    corresponding : yes    # Define only one corresponding author
    email         : "richard.ramsey@hest.ethz.ch"
    role:
      - "Conceptualization"
      - "Visualisation"
      - "Supervision"
      - "Project Administration"
      - "Validation"
      - "Writing - Review & Editing"
affiliation:
  - id            : "1"
    institution   : "Macquarie University"
  - id            : "2"
    institution   : "ETH Zürich"
authornote: |
  Samantha Parker. School of Psychological Sciences, Macquarie University.
  Emily S. Cross. Social Brain Sciences Lab, Department of Humanities, Social and Political Sciences (D-GESS), ETH Zürich.
  Richard Ramsey. Neural Control of Movement Lab, Department of Health Sciences and Technology (D-HEST) and Social Brain Sciences Lab, Department of Humanities, Social and Political Sciences (D-GESS), ETH Zürich.
  <!-- Enter author note here. -->
abstract: |
  Evidence accumulation modelling has been shown to uncover new insights into the cognitive mechanisms that underlie decision making from behavioural data. By jointly modelling reaction time and accuracy data, such decision models estimate latent variables that represent distinct computaitonal processes, such as stimulus encoding, response caution and the quality of information processing. In this study we use an evidence accumulation model, the Linear Ballistic Accumulator (LBA), to shed new light into the mechanisms that underlie learning based on linguistic and action-based training. The LBA model is applied to behavioural data from a previously published training study where participants learn to name, tie or name and tie a set of knots. Our results show that training is multifaceted and associated with an increase in stimulus-encoding time, a reduction in response caution, as well as an increase in the speed of information accumulation. Furthermore, the results showed that there was an added benefit to the rate of evidence accumulation when naming and tying experience were combined. This latter finding suggests that performance benefits from multi-modal training may be instantiated in computational processes that are associated with the quantity and quality of information accumulation during decision making. Overall, in applying this computational approach to accuracy and reaction time data, we uncover new insights into the mechanisms that govern experience-dependent plasticity.

>>>>

---
title             : "Evidence accumulation modelling offers new insights into the cognitive mechanisms that underlie linguistic and action-based training"
shorttitle        : "Decision models and experience-dependent plasticity"
author: 
  - name          : "Samantha Parker"
    affiliation   : "1"
    # corresponding : yes    # Define only one corresponding author
    # address       : "School of Psychological Sciences, 16 University Avenue, Macquarie University, NSW 2109"
    # email         : "samantha.parker@students.mq.edu.au"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Data curation"
      - "Formal analysis"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Emily S. Cross"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Resources"
      - "Writing - Review & Editing"
  - name          : "Richard Ramsey"
    affiliation   : "2"
    corresponding : yes    # Define only one corresponding author
    address       : "ETH Zurich"
    email         : "richard.ramsey@hest.ethz.ch"
    role:
      - "Conceptualization"
      - "Visualisation"
      - "Supervision"
      - "Project Administration"
      - "Validation"
      - "Writing - Review & Editing"
affiliation:
  - id            : "1"
    institution   : "Macquarie University"
  - id            : "2"
    institution   : "ETH Zürich"
authornote: |
  Samantha Parker. School of Psychological Sciences, Macquarie University.
  Emily S. Cross. Social Brain Sciences Lab, Department of Humanities, Social and Political Sciences (D-GESS), ETH Zürich.
  Richard Ramsey. Neural Control of Movement Lab, Department of Health Sciences and Technology (D-HEST) and Social Brain Sciences Lab, Department of Humanities, Social and Political Sciences (D-GESS), ETH Zürich.
  <!-- Enter author note here. -->
abstract: |
  Evidence accumulation modelling has been shown to uncover new insights into the cognitive mechanisms that underlie decision making from behavioural data. By jointly modelling reaction time and accuracy data, such decision models estimate latent variables that represent distinct computaitonal processes, such as stimulus encoding, response caution and the quality of information processing. In this study we use an evidence accumulation model, the Linear Ballistic Accumulator (LBA), to shed new light into the mechanisms that underlie learning based on linguistic and action-based training. The LBA model is applied to behavioural data from a previously published training study where participants learn to name, tie or name and tie a set of knots. Our results show that training is multifaceted and associated with an increase in stimulus-encoding time, a reduction in response caution, as well as an increase in the speed of information accumulation. Furthermore, the results showed that there was an added benefit to the rate of evidence accumulation when naming and tying experience were combined. This latter finding suggests that performance benefits from multi-modal training may be instantiated in computational processes that are associated with the quantity and quality of information accumulation during decision making. Overall, in applying this computational approach to accuracy and reaction time data, we uncover new insights into the mechanisms that govern experience-dependent plasticity.
